```{r setup}
#| echo: false
#| message: false
#| warning: false
set.seed(506)
```
## Import Packages
```{r}
#install.packages('fpp2')
#install.packages('import')
#install.packages('caret')
#install.packages('readr')
#install.packages('tidyverse')
#install.packages('zoo')
#install.packages('gridExtra')
#install.packages('forecast')

library(caret)
library(forecast)
library(gridExtra)
#install.packages('import')
library(import)
library(readr)
library(fpp2)
library(zoo)
library(tidyverse)
```
## Load Data
```{r}
# read in data
df <- read_csv('PLOO_water_temperature_2022_datasd.csv', show_col_types = FALSE)
head(df)

# convert the start date string into a POSIXct object
df$datetime_pst <- as.POSIXct(df$datetime_pst, format='%Y-%m-%d %H:%M:%S')
head(df)
```
## Split Data by depth
```{r}
## splice data for each depth ##
df_1m <- subset(df, depth_m==1)
df_9m <- subset(df, depth_m==9)
df_20m <- subset(df, depth_m==20)
df_30m <- subset(df, depth_m==30)
df_45m <- subset(df, depth_m==45)
df_60m <- subset(df, depth_m==60)
df_74m <- subset(df, depth_m==74)
df_87m <- subset(df, depth_m==87)

#Create empty data frame
output = data.frame()

# add the summary table for each depth to the output df
depths_df <- list(df_1m, df_9m, df_20m, df_30m, df_45m, df_60m, df_74m, df_87m)
for (x in depths_df) {
  sum = summary(x$value)
  output = rbind(output,sum)
}

# name columns
colnames(output) = c("Min.","1st Qu.","Median","Mean","3rd Qu.","Max.", "NA's")

# add a column to label depth
depths <- c(1, 9, 20, 30, 45, 60, 74, 87)
output$Depth <- depths
output <- output[c("Depth","Min.","1st Qu.","Median","Mean","3rd Qu.","Max.", "NA's")]
output

for (x in depths_df) {print(sum(is.na(x)))}   

impute_null_with_average <- function(x, y) {
  if (is.na(x)) {
    previous_index <- which(!is.na(y))[1]
    next_index <- which(!is.na(y))[2]
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- y[previous_index]
      next_value <- y[next_index]
      return((previous_value + next_value) / 2) 
    }
  }
  return(x) 
}
```
## Null values replaced by Imputing
```{r}
# apply the function to impute null values
df_1m <- df_1m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_1m$value))
df_9m <- df_9m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_9m$value))
df_20m <- df_20m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_20m$value))
df_30m <- df_30m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_30m$value))
df_45m <- df_45m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_45m$value))
df_60m <- df_60m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_60m$value))
df_74m <- df_74m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_74m$value))
df_87m <- df_87m %>% rowwise() %>% mutate(value = impute_null_with_average(value, df_87m$value))

for (y in depths_df) {print(sum(is.na(df_9m)))}
```
## Time Series for each Depth
```{r}
# Create time series object
df_1m_ts <- ts(df_1m$value)
df_9m_ts <- ts(df_9m$value)
df_20m_ts <- ts(df_20m$value)
df_30m_ts <- ts(df_30m$value)
df_45m_ts <- ts(df_45m$value)
df_60m_ts <- ts(df_60m$value)
df_74m_ts <- ts(df_74m$value)
df_87m_ts <- ts(df_87m$value)

# plot all depths
autoplot(df_1m_ts, main='Real-time Oceanographic Mooring System: Point Loma Ocean 2022', 
    series='1m depth', ylab='Water Temperature (Degrees Celsius)') + 
    autolayer(df_9m_ts, series='9m depth') +
    autolayer(df_20m_ts, series='20m depth') +
    autolayer(df_30m_ts, series='30m depth') +
    autolayer(df_45m_ts, series='45m depth') +
    autolayer(df_60m_ts, series='60m depth') +
    autolayer(df_74m_ts, series='74m depth') +
    autolayer(df_87m_ts, series='87m depth')

depths_ts <- list(df_1m_ts, df_9m_ts, 
                  df_20m_ts, df_30m_ts, 
                  df_45m_ts, df_60m_ts, 
                  df_74m_ts, df_87m_ts)

n = 1
myplots <- list()
for (x in depths_ts) {
    y <- autoplot(x, main=paste0('RTOMS Readings at ', 
        depths[[n]] ,'m'), ylab='Water Temperature (Degrees Celsius)')
    myplots[[n]] <- y
    n = n +1
}

len <- length(myplots)
nCol <- floor(sqrt(len))
do.call("grid.arrange", c(myplots, ncol=nCol))

# see distribution of water temperatures
par(mfrow=c(2,4))
hist(df_1m$value, main='1m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_9m$value, main='9m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_20m$value, main='20m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_30m$value, main='30m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_45m$value, main='45m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_60m$value, main='60m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_74m$value, main='74m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_87m$value, main='87m', xlab= 'Water Temperature (Degrees Celsius)')
```
## Differencing
```{r}
df_1m_diff <- diff(df_1m_ts, frequency=365,start=c(2022,1,1))
df_9m_diff <- diff(df_9m_ts, frequency=365,start=c(2022,1,1))
df_20m_diff <- diff(df_20m_ts, frequency=365,start=c(2022,1,1))
df_30m_diff <- diff(df_30m_ts, frequency=365,start=c(2022,1,1))
df_45m_diff <- diff(df_45m_ts, frequency=365,start=c(2022,1,1))
df_60m_diff <- diff(df_60m_ts, frequency=365,start=c(2022,1,1))
df_74m_diff <- diff(df_74m_ts, frequency=365,start=c(2022,1,1))
df_87m_diff <- diff(df_87m_ts, frequency=365,start=c(2022,1,1))

#Create empty data frame
diff_df = data.frame()

# add the summary table for each depth to the output df
diff_dfs <- list(df_1m_diff, df_9m_diff, df_20m_diff, df_30m_diff, df_45m_diff, df_60m_diff, df_74m_diff, df_87m_diff)
for (x in diff_dfs) {
  sum = summary(x)
  diff_df = rbind(diff_df,sum)
}

# name columns
colnames(diff_df) = c("Min.","1st Qu.","Median","Mean","3rd Qu.","Max.")

# add a column to label depth
diff_df$Depth <- depths
diff_df <- diff_df[c("Depth","Min.","1st Qu.","Median","Mean","3rd Qu.","Max.")]
diff_df

par(mfrow=c(2, 4))
# the other fxn to make a nice grid doesn't work with the acf graphs
for (z in diff_dfs) {
    acf(z, main = "ACF of Differenced Time Series")
}

par(mfrow=c(2, 4))
for (z in diff_dfs) {    
    pacf(z, main = "PACF of Differenced Time Series")
}
```
## Daily Averages
```{r}
## 9m ##
daily_avg_1m <- df_1m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_9m <- df_9m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_20m <- df_20m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_30m <- df_30m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_45m <- df_45m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_60m <- df_60m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_74m <- df_74m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))
daily_avg_87m <- df_87m %>% group_by(date=as.Date(datetime_pst)) %>% 
    summarize(avg_value = mean(value, na.rm=TRUE))

daily_avg_1m_ts <- ts(daily_avg_1m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_9m_ts <- ts(daily_avg_9m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_20m_ts <- ts(daily_avg_20m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_30m_ts <- ts(daily_avg_30m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_45m_ts <- ts(daily_avg_45m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_60m_ts <- ts(daily_avg_60m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_74m_ts <- ts(daily_avg_74m$avg_value, frequency=365, start=c(2022,1,1))
daily_avg_87m_ts <- ts(daily_avg_87m$avg_value, frequency=365, start=c(2022,1,1))

average_ts <- list(daily_avg_1m_ts, daily_avg_9m_ts, 
                   daily_avg_20m_ts, daily_avg_30m_ts, 
                   daily_avg_45m_ts, daily_avg_60m_ts, 
                   daily_avg_74m_ts, daily_avg_87m_ts)

n = 1
myplots_avg <- list()
for (x in average_ts) {
    y <- autoplot(x, main=paste0('Daily Average at ', depths[[n]] ,'m'), 
                  ylab='Water Temperature (Degrees Celsius)')
    myplots_avg[[n]] <- y
    n = n +1
}

len <- length(myplots_avg)
nCol <- floor(sqrt(len))
do.call("grid.arrange", c(myplots_avg, ncol=nCol))

# Export daily average measurements at 45m for visualizations in Tableau
write_csv(daily_avg_45m, "ploo_daily_avg_45m.csv")
```
## Visualize the chosen time series
```{r}
chosen_ts <- daily_avg_45m_ts
autoplot(chosen_ts, main=paste0('Daily Average at ', 45 ,'m'), ylab='Water Temperature (Degrees Celsius)')
```
## Partition
```{r}
# Set the number of validation points
nValid <- 36

# Calculate the number of training points
nTrain <- length(daily_avg_45m_ts) - nValid

# Create training and validation sets
train_ts <- window(daily_avg_45m_ts, end = c(2022, nTrain))
valid_ts <- window(daily_avg_45m_ts, start = c(2022, nTrain + 1), end = c(2022, nTrain + nValid))

#daily_avg_45m_ts
length(train_ts)
length(valid_ts)
```
# Train Models
## 6 Total Models will be trained
## Naive, Auto-Arima, Arima, Neural Network, ETS, and Linear Regression
```{r}
# Naive
naive_model <- naive(train_ts)
summary(naive_model)

# Auto-ARIMA
auto_arima_model <- auto.arima(train_ts)
summary(auto_arima_model)

autoplot(auto_arima_model$fitted, series='Auto-ARIMA', alpha=0.4) + autolayer(train_ts, series='Training Data', color='black') + autolayer(naive_model$fitted, series='Naive Model', alpha=0.4)
```
# ARIMA
## Look at ACF and PACF plots to determine ARIMA components
```{r}
acf(train_ts)

## All ACF values are significant. Take the first difference. 
acf(diff(train_ts))
```
## Taking the first difference removed most of the auto-correlation in the series. 
```{r}
pacf(train_ts)
pacf(diff(train_ts))


arima_model <- arima(train_ts, order=c(4,1,1))
summary(arima_model)
```

# Other models
```{r}
# Neural Network
nn_model <- nnetar(train_ts, repeats = 20, p = 7, P = 0, size = 5)
summary(nn_model)



# ETS
ets_model <- ets(train_ts)
summary(ets_model)


# Linear Regression
lr_model <- tslm(train_ts ~ trend + I(trend^2))
summary(lr_model)

```


# FOR ALL MODELS

```{r}

list_of_models <- list(naive_model, auto_arima_model, arima_model, nn_model, ets_model, lr_model)
list_of_models_names <- c('Naive', 'Auto_ARIMA', 'Arima', 'Neural Network', 'ETS', 'LR')

```
## Graph the residuals of each model against the training set
```{r}


n = 1
resid <- list()
for (x in list_of_models) {
    y <- autoplot(x$residuals)
    resid[[n]] <- y
    n = n +1
}

len <- length(resid)
nCol <- floor(sqrt(len))
do.call("grid.arrange", c(resid, ncol=nCol))

```

# Perfromance Metrics
```{r}

# Create Predictions
naive_pred<- forecast(auto_arima_model, h = nValid)
auto_arima_pred <- forecast(auto_arima_model, h = nValid)
arima_pred <- forecast(arima_model, h = nValid)
ets_pred <- forecast(ets_model, h = nValid)
nn_pred <- forecast(nn_model, h = nValid)
lr_pred <- forecast(lr_model, h = nValid)
```
# Create a table with the RMSE for each model
```{r}
#Create empty data frame
errors = data.frame()
# predictions
list_of_model_pred <- list(naive_pred, auto_arima_pred, arima_pred, nn_pred, ets_pred, lr_pred)
for (x in list_of_model_pred) {
  sum = accuracy(x,valid_ts)[,2]
  errors = rbind(errors,sum)
}

# add a cxolumn to label models
errors$model <- list_of_models_names

# name columns
colnames(errors) = c("RMSETraining", "RMSETest", "Model")


errors <- errors[c("Model", "RMSETraining", "RMSETest")]
errors

```
# Graph RMSE
```{r}
#par(mfrow=c(2, 2))
# Bar chart for Training RMSE
barplot(errors$RMSETraining, names.arg = list_of_models_names, col = 'lightblue',
        main = 'Training RMSE per Model', xlab = 'Model', ylab = 'RMSE')

# Bar chart for Test RMSE
barplot(errors$RMSETest, names.arg = list_of_models_names, col = 'lightgreen',
        main = 'Test RMSE per Model', xlab = 'Model', ylab = 'RMSE')


```
# graph Naive Forecast predictions
```{r}

# Plotting
plot(naive_pred, ylab = "Daily Average", xlab = "Time", bty = "l", xaxt = "n", main = "Naive Forecast", flty = 2)
lines(valid_ts, col = 'red')

```

# Other Forecast Plots

```{r}
## ETS

plot(ets_pred, main = "ETS Model Forecast", ylab = "value", xlab = "time")
lines(valid_ts, col = 'red')



# LR

# Plotting
plot(lr_pred, ylim = c(min(daily_avg_45m_ts), max(daily_avg_45m_ts)), ylab = "Daily Average", xlab = "Time", bty = "l",
     xaxt = "n", main = "", flty = 2)
lines(lr_model$fitted, lwd = 2)
axis(1, at = time(valid_ts), labels = format(time(valid_ts)))
lines(valid_ts, col = 'red')





