
### Library and data import
```{r setup}
#| echo: false
#| message: false
#| warning: false

# library import
library(readr)
library(fpp2)
library(zoo)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(forecast)

set.seed(506)
```

```{r}
# read in data
df <- read_csv('PLOO_water_temperature_2022_datasd.csv', show_col_types = FALSE)

# convert the start date string into a POSIXct object
df$datetime_pst <- as.POSIXct(df$datetime_pst, format='%Y-%m-%d %H:%M:%S')
```

### Data splicing
All temperature readings will be divided into different dataframes according to depth
```{r}
## splice data for each depth ##
df_1m <- subset(df, depth_m==1)
df_9m <- subset(df, depth_m==9)
df_20m <- subset(df, depth_m==20)
df_30m <- subset(df, depth_m==30)
df_45m <- subset(df, depth_m==45)
df_60m <- subset(df, depth_m==60)
df_74m <- subset(df, depth_m==74)
df_87m <- subset(df, depth_m==87)

# statistics
summary(df_1m$value)
summary(df_9m$value)
summary(df_20m$value)
summary(df_30m$value)
summary(df_45m$value)
summary(df_60m$value)
summary(df_74m$value)
summary(df_87m$value)
```

Consider measurements at 9m, 20m, 45m, 60m, or 87m for analysis since they have the fewest nulls. 

### Null imputation
Replace nulls with average value of previous and next temperature readings. If both previous and next values are NA, loop through until the next non-NA value is found. 
```{r}
## 1 meter ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_1m$value))[1]
    next_index <- which(!is.na(df_1m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_1m$value[previous_index]
      next_value <- df_1m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_1m <- df_1m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_1m))

## 9 meter ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_9m$value))[1]
    next_index <- which(!is.na(df_9m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_9m$value[previous_index]
      next_value <- df_9m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_9m <- df_9m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_9m))


## 20m ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_20m$value))[1]
    next_index <- which(!is.na(df_20m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_20m$value[previous_index]
      next_value <- df_20m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_20m <- df_20m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_20m))

## 30m ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_30m$value))[1]
    next_index <- which(!is.na(df_30m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_30m$value[previous_index]
      next_value <- df_30m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_30m <- df_30m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_30m))

## 45m ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_45m$value))[1]
    next_index <- which(!is.na(df_45m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_45m$value[previous_index]
      next_value <- df_45m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_45m <- df_45m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_45m))

## 60m ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_60m$value))[1]
    next_index <- which(!is.na(df_60m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_60m$value[previous_index]
      next_value <- df_60m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_60m <- df_60m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_60m))

## 74m ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_74m$value))[1]
    next_index <- which(!is.na(df_74m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_74m$value[previous_index]
      next_value <- df_74m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_74m <- df_74m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_74m))

## 87m ##
# create function to use next available non-NA values for previous and next rows
impute_null_with_average <- function(x) {
  if (is.na(x)) {
    previous_index <- which(!is.na(df_87m$value))[1]
    next_index <- which(!is.na(df_87m$value))[2]
    
    if (!is.na(previous_index) && !is.na(next_index)) {
      previous_value <- df_87m$value[previous_index]
      next_value <- df_87m$value[next_index]
      return((previous_value + next_value) / 2)
    }
  }
  
  return(x)
}

# apply the function to impute null values
df_87m <- df_87m %>%
  rowwise() %>%
  mutate(value = impute_null_with_average(value))

# show null counts
sum(is.na(df_87m))
```

### Create time series objects and visualize
```{r}
# Create time series object
df_1m_ts <- ts(df_1m$value)
df_9m_ts <- ts(df_9m$value)
df_20m_ts <- ts(df_20m$value)
df_30m_ts <- ts(df_30m$value)
df_45m_ts <- ts(df_45m$value)
df_60m_ts <- ts(df_60m$value)
df_74m_ts <- ts(df_74m$value)
df_87m_ts <- ts(df_87m$value)

# plot all depths
autoplot(df_1m_ts, main='Real-time Oceanographic Mooring System: Point Loma Ocean 2022', series='1m depth',
ylab='Water Temperature (Degrees Celsius)') + 
  autolayer(df_9m_ts, series='9m depth') +
  autolayer(df_20m_ts, series='20m depth') +
  autolayer(df_30m_ts, series='30m depth') +
  autolayer(df_45m_ts, series='45m depth') +
  autolayer(df_60m_ts, series='60m depth') +
  autolayer(df_74m_ts, series='74m depth') +
  autolayer(df_87m_ts, series='87m depth')

## visualize depths of interest ##
# 1m
autoplot(df_1m_ts, main='RTOMS Readings at 1m', ylab='Water Temperature (Degrees Celsius)')
# 20m
autoplot(df_20m_ts, main='RTOMS Readings at 20m', ylab='Water Temperature (Degrees Celsius)')
# 45m
autoplot(df_45m_ts, main='RTOMS Readings at 45m', ylab='Water Temperature (Degrees Celsius)')
# 60m
autoplot(df_60m_ts, main='RTOMS Readings at 60m', ylab='Water Temperature (Degrees Celsius)')
# 87m
autoplot(df_87m_ts, main='RTOMS Readings at 87m', ylab='Water Temperature (Degrees Celsius)')

# see distribution of water temperatures
par(mfrow=c(2,4))
hist(df_1m$value, main='1m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_9m$value, main='9m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_20m$value, main='20m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_30m$value, main='30m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_45m$value, main='45m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_60m$value, main='60m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_74m$value, main='74m', xlab= 'Water Temperature (Degrees Celsius)')
hist(df_87m$value, main='87m', xlab= 'Water Temperature (Degrees Celsius)')
```

### Differencing
```{r}
df_9m_diff <- diff(df_9m_ts, lag=6)

# Plotting the results
autoplot(df_9m_diff) +
  labs(title = "Differenced Time Series (depth_m = 9)", y = "Difference") +
  theme_classic()
```
```{r}
# Fit ARIMA model with differencing
df_9m_diff_model <- Arima(df_9m_ts, order = c(1, 1, 1), seasonal = list(order = c(1, 1, 1)))
df_9m_ARIMA <- residuals(df_9m_diff_model)
plot(df_9m_ARIMA , main = "Differenced Time Series", ylab = "Differenced Values")
```


```{r}
# Fit Logistic Regression model
df_9m_train <- window(df_9m_ts, end = length(df_9m_ts) - 24)
df_9m_test <- window(df_9m_ts, start = length(df_9m_ts) - 23)

# Fit ETS model to the differenced training set
df_9m_model <- ets(df_9m_train)

# Generate forecasts for the differenced test set
df_9m_pred <- forecast(df_9m_model, h = 24)

summary(df_9m_model)
```
```{r}
plot(df_9m_ts, col = "blue", main = "Time Series Forecasting with Logistic Regression Model")
lines(df_9m_train, col = "red", lty = 2, lwd = 2)
lines(df_9m_pred$mean, col = "green", lty = 2, lwd = 2)

# Add legend
legend("topleft", legend = c("Original Time Series", "Training Set", "Forecast"), col = c("blue", "red", "green"), lty = c(1, 2, 2))

```








